### 2010-10-13 papers/BMD2010/PhysicalParametersPaper/paper.tex
cleaning up the paper, combined some subfigures

\begin{center}
\begin{tabular}{c}
$^*$ Mechanical and Aerospace Engineering\\
University of California, Davis\\
One Shields Avenue, Davis, CA, 95616, USA\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu\\
\end{tabular}
\begin{tabular}{c}
$^\dag$ Laboratory for Engineering Mechanics\\
Delft University of Technology\\
Mekelweg 2, 2628CD Delft, The Netherlands\\
e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com\\
\end{tabular} \\ \vspace{2.5ex}

\section*{ABSTRACT}
Accurate measurements of a bicycle's physical parameters are required for
realistic dynamic simulations and analysis. For the most basic models the
geometry, mass, mass location and mass distributions must be measured. More complex models
require measurements of tire characteristics, human characteristics, friction, stiffness, damping, etc. This
paper concerns the measurement of the minimal bicycle parameters required for
the benchmark bicycle presented in~\cite{Meijaard2007}. This
model is composed of four rigid bodies, has ideal rolling and frictionless joints,
and is assumed to be laterally symmetric. A set of 25
parameters is used to describe the geometry, mass, mass location and
mass distribution of each of the rigid bodies. The experimental methods
described herein are based primarily on the work
done in~\cite{Kooijman2006} but have been refined for improved accuracy and
methodology. Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
similar techniques with scooters. We measured the characteristics of six
different bicycles, two of which were set up in two different configurations.
This is a total of eight different parameter sets that can be used with, but not
limited to, the benchmark bicycle model. The accuracy of all the measurements
are presented up through the eigenvalue prediction of the linear model and are
based on error propagation theory with correlations taken into account.

@@

@@

\subsection{Wheelbase}
The wheelbase was measured directly using a tape measure while the bicycle was
in the same upright fixed position used when measuring the head tube angle.
\section{Mass}

@@

resolution of 100g. Then each of the four bicycle parts were measured using a
Molen 20kg scale with a resolution of 20g.

@@

\section{Center of Mass}
\subsection{Wheels}
The centers of mass of the wheels are assumed to be at their geometrical centers.
\subsection{Rear frame}
The rear frame was hung in three orientations as a torsional pendulum (the
pendulum was used to measure moments of inertia). I
assumed that the frame was laterally symmetric. The
frame could rotate about a joint such that gravity aligned the center of mass
with the pendulum axis. The orientation angle of the headtube $\alpha$
Figure~\ref{fig:angles} relative
to the earth was
measured using the digital level Figure~\ref{fig:level}. A string was aligned with the pendulum axis

@@

axle and the string (CoM line) was measured by aligning a ruler perpendicular to
the string. The distance $a$ was negative if the string fell to the right of

@@

@@

		\includegraphics[width=4in]{../../../images/level.jpg}

@@

The frame rotation angle $\beta$ is defined as rotation of the frame in the

@@

		\includegraphics[width=4in]{../../../figures/angles.pdf}
	\caption{Pictorial description of the angles and dimensions that related
    $XYZ_P$.}
	\label{fig:angles}

@@

The pendulum axis $X_P$ is simply a line in the nominal bicycle reference frame with a slope $m$ and a z-intercept $b$. The slope can be shown to be

@@

   y = m_ix+b_i

@@

The CoM location can be calculated by finding the intersection of these three

@@

leads me to think that the center of mass is located at the centroid of the

@@

        \includegraphics{../../../figures/triangle.pdf}
    \caption{Exaggerated intersection of the three pendulum axes and the
    \label{fig:triangle}

@@

>>>>

\begin{center}
\begin{tabular}{c}
$^*$ Mechanical and Aerospace Engineering\\
University of California, Davis\\
One Shields Avenue, Davis, CA, 95616, USA\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu\\
\end{tabular}
\begin{tabular}{c}
$^\dag$ Laboratory for Engineering Mechanics\\
Delft University of Technology\\
Mekelweg 2, 2628CD Delft, The Netherlands\\
e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com\\
\end{tabular} \\ \vspace{2.5ex}
\end{center}
\section*{ABSTRACT}
Accurate measurements of a bicycle's physical parameters are required for
realistic dynamic simulations and analysis. For the most basic models the
geometry, mass, mass location and mass distributions must be measured. More complex models
require measurements of tire characteristics, human characteristics, friction, stiffness, damping, etc. This
paper concerns the measurement of the minimal bicycle parameters required for
the benchmark bicycle presented in~\cite{Meijaard2007}. This
model is composed of four rigid bodies, has ideal rolling and frictionless joints,
and is assumed to be laterally symmetric. A set of 25
parameters is used to describe the geometry, mass, mass location and
mass distribution of each of the rigid bodies. The experimental methods
described herein are based primarily on the work
done in~\cite{Kooijman2006} but have been refined for improved accuracy and
methodology. Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
similar techniques with scooters. We measured the characteristics of six
different bicycles, two of which were set up in two different configurations.
This is a total of eight different parameter sets that can be used with, but not
limited to, the benchmark bicycle model. The accuracy of all the measurements
are presented up through the eigenvalue prediction of the linear model and are
based on error propagation theory with correlations taken into account.

@@

\label{sec:headtube}

@@

\subsection{WHEELBASE}
We measured the wheelbase when the bicycle was in the same nominal
configuration as described in Sec.~\ref{sec:headtube}. We used a tape measure
to measure the distance from one wheel axle center to the other with a 0.002 m
accuracy.
\section{MASS}

@@

resolution of 100 grams. The total mass was only used for comparison purposes.
Each of the four bicycle parts were measured using a Molen 20 kilogram scale with a
resolution of 20 grams. The accuracy was conservatively assumed to also be
$\pm20$ grams.

@@

\section{CENTER OF MASS}
\subsection{WHEELS}
The centers of mass of the wheels are assumed to be at their geometrical
centers to comply with the Whipple model.
\subsection{REAR FRAME}
The rear frame was hung in three orientations as a torsional pendulum (both for
the center of mass measurements and the moment of inertia measurements
described in Sec.~\ref{sec:moi}). We assumed that the frame was laterally
symmetric, complying with the Whipple model. The frame could rotate about a joint
such that gravity aligned the center of mass with the pendulum axis. The
orientation angle of the headtube, $\alpha_\mathrm{B}$, Fig.~\ref{fig:angles} relative
to the earth was measured using a digital level ($\pm0.2^{\circ}$ accuracy),
Figure~\ref{fig:level}. A string was aligned with the pendulum axis

@@

axle and the string was measured by aligning a ruler perpendicular to
the string. The distance $a_\mathrm{B}$ was negative if the string fell to the right of

@@

    \subfloat[]{\label{fig:angles}\includegraphics[width=3in]{../../../figures/angles.pdf}}
    \subfloat[]{\label{fig:triangle}\includegraphics[width=3in]{../../../figures/triangle.pdf}}
    \caption{\subref{fig:angles} Pictorial description of the angles and dimensions that related
    $XYZ_P$. \subref{fig:triangle} Exaggerated intersection of the three pendulum axes and the

@@

		\includegraphics[width=3in]{../../../images/level.jpg}

@@

The frame rotation angle $\beta_\mathrm{B}$ is defined as rotation of the frame in the

@@

@@

The center of mass can be found by realizing that the pendulum axis $X_P$ is
simply a line in the nominal bicycle reference frame with a slope $m$ and a
z-intercept $b$ where the $i$ subscript corresponds the different frame
orientations~\ref{fig:triangle}. The slope can be shown to be

@@

   z = m_ix+b_i

@@

The mass center location can be calculated by finding the intersection of these three

@@

lead us to think that the center of mass is located at the centroid of the

@@

@@

\label{sec:moi}
